---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---

Clean up memory and load needed libraries:
```{r}
    ## Clean up
    rm(list=(ls()))

    ## Load libraries
    library(dplyr)
```


## Loading and preprocessing the data

Load data and group by date:
```{r}
## Load data/display summary
data <- read.csv("data\\activity.csv")
summary(data)
```


## What is mean total number of steps taken per day?

Analyze daily data ignoring NAs:
```{r}
## Group data by date
data.group.date <- group_by(data, date)

## Sum # of steps per day
data.sum.date <- summarize(data.group.date, steps = sum(steps))

## Display histogram
hist(data.sum.date$steps, col = "blue", xlab = "Steps per day", main = "Total Number of Steps Taken Each Day")

## Report mean and median # of steps
summarize(data.sum.date, mean = mean(steps, na.rm = TRUE), median = median(steps, na.rm = TRUE))
```


## What is the average daily activity pattern?

Analyze average daily activity pattern (ignoring NAs):
```{r}
## Group data by interval
data.group.interval <- group_by(data, interval)

## Average # of steps per interval
data.sum.interval <- summarize(data.group.interval, mean = mean(steps, na.rm = TRUE))

## Display time series
plot(data.sum.interval$interval, data.sum.interval$mean, type = "l", xlab = "Interval", ylab = "Steps", main = "Average # of Steps Taken (averaged across all days)")

## Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps
data.sum.interval[which(data.sum.interval$mean == max(data.sum.interval$mean)), ]
```


## Imputing missing values

Determine impact of missing values. We then imput the missing values by using the mean of non-NA observations.
```{r}
## Calculate the toal number of missing values in the dataset
sum(is.na(data$steps))

## Determine value to use for missing values (mean for non-NA)
na_fill_value <- mean(data$steps, na.rm = TRUE)

## Create a new dataset that is equal to the original dataset but with the missings data filled in
data.nona <- data
data.nona[is.na(data.nona)] <- na_fill_value

## Group data by date
data.nona.group.date <- group_by(data.nona, date)

## Sum # of steps per day
data.nona.sum.date <- summarize(data.nona.group.date, steps = sum(steps))

## Display histogram
hist(data.nona.sum.date$steps, col = "cyan", xlab = "Steps per day", main = "Total Number of Steps Taken Each Day")

## Report mean and median # of steps
summarize(data.nona.sum.date, mean = mean(steps, na.rm = TRUE), median = median(steps, na.rm = TRUE))

```


## Are there differences in activity patterns between weekdays and weekends?
